\LARGE \bf Laser Tracker Placement Optimization for highly flexible manufacturing systems

### 2024-07-24 paper/paper.tex
added deviation cost and laser tracker

\section{Formulation of the optimization problem}
To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
Similar to approaches such as \cite{ieee_sensors} we divide the visibility into two components.
Line of sight visibility and field of view visibility.
Both  are depicted in figure \ref{fig:visibility}.
\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
        \caption{Field of view (VOW) angle range and line of sight (LOS) components of marker visibility}
        \label{fig:visibility}
\end{figure}
Mathematically we can define line of sight visibility (los visibility) as the following binary function:
\begin{equation}
    f(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
Note that $p_m$ is typically dependent on time as it is attached to a moving manufacuting system such as a robot.
The field of view objective can similarly be defined as a binary function:
\begin{equation}
    g(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
However in practice it might make sense to not be close to the boundary of the field of view.
Instead one can define a continuous function that is 1 in the center of the field of view and 0 at the boundary.
The simplest way to encode this using the angle $\alpha$ between the normal of the marker and the line of sight of the laser tracker.
\begin{equation}
    g(p_m,p_l) = \cos(\alpha)
\end{equation}
The final objective function is then a weighted sum of the two components:
\begin{equation}
    h(p_m,p_l) = \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l)
    \label{eq:objective}
\end{equation}
where $\lambda$ is a weighting factor between 0 and 1 that can be adjusted to prioritize one component over the other.
After defining the objective function we can now define the design variables and constraints.
The design variables are the pose of the laser tracker $p_l$ and the pose of the marker $p_m$.
However the pose of these objects is not arbitrary.
For the purposes of this paper we assume that the marker is attached to the flange of the manufacturing system.
This means that we only define a Transformation matrix $T_{m}$ that describes the pose of the marker relative to the flange.
The final pose of the marker $p_m$ is then given by the flange transformation $T_{f}(q)$ where $q$ is the joint configuration of the robot as well as the transformation $T_{m}$:
\begin{equation}
    p_m = T_{f}(q)T_{m}
\end{equation}
While the matrix $T_m$ is a 4x4 matrix, we can parameterize it using the position of the marker $p_{p_m}$ and the orientation of the marker $p_{o_m}$.
This results in a 6 dimensional design space for the marker pose.
This space is further constraint to avoid any unrealistic marker positions (such as 10 meters above the robot).
Ideally the marer should be attached close to the surface of the manufacutring system to avoid any unneccessarily large constructions that might collide with the environment or cause vibrations.
This can be encoded using a cylindrical constraint on the position of the marker:
\begin{equation}
        \begin{split}
    \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
        0 \leq p_{p_m,z} &\leq h
        \end{split}
\end{equation}
where $r$ is the radius of the cylinder and $h$ is the height of the cylinder.
The marker of course also cant be within the mesh of the manufacturing system, but since this would result in a line of sight visibility of 0 we do not need to encode this constraint explicitly.

>>>>

\section{Formulation of the optimization problem}
To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
Similar to approaches such as \cite{ieee_sensors} we divide the visibility into two components.
Line of sight visibility and field of view visibility.
Both  are depicted in figure \ref{fig:visibility}.
\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
        \caption{Field of view (VOW) angle range and line of sight (LOS) components of marker visibility}
        \label{fig:visibility}
\end{figure}
Mathematically we can define line of sight visibility (los visibility) as the following binary function:
\begin{equation}
    f(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
Note that $p_m$ is typically dependent on time as it is attached to a moving manufacuting system such as a robot.
The field of view objective can similarly be defined as a binary function:
\begin{equation}
    g(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
However in practice it might make sense to not be close to the boundary of the field of view.
Instead one can define a continuous function that is 1 in the center of the field of view and 0 at the boundary.
The simplest way to encode this using the angle $\alpha$ between the normal of the marker and the line of sight of the laser tracker.
\begin{equation}
    g(p_m,p_l) = \cos(\alpha)
\end{equation}
The final objective function is then a weighted sum of the two components:
\begin{equation}
    h(p_m,p_l) = \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l)
    \label{eq:objective}
\end{equation}
where $\lambda$ is a weighting factor between 0 and 1 that can be adjusted to prioritize one component over the other.
After defining the objective function we can now define the design variables and constraints.
The design variables are the pose of the laser tracker $p_l$ and the pose of the marker $p_m$.
However the pose of these objects is not arbitrary.
For the purposes of this paper we assume that the marker is attached to the flange of the manufacturing system.
This means that we only define a Transformation matrix $T_{m}$ that describes the pose of the marker relative to the flange.
The final pose of the marker $p_m$ is then given by the flange transformation $T_{f}(q)$ where $q$ is the joint configuration of the robot as well as the transformation $T_{m}$:
\begin{equation}
    p_m = T_{f}(q)T_{m}
\end{equation}
While the matrix $T_m$ is a 4x4 matrix, we can parameterize it using the position of the marker $p_{p_m}$ and the orientation of the marker $p_{o_m}$.
This results in a 6 dimensional design space for the marker pose.
This space is further constraint to avoid any unrealistic marker positions (such as 10 meters above the robot).
Ideally the marer should be attached close to the surface of the manufacutring system to avoid any unneccessarily large constructions that might collide with the environment or cause vibrations.
This can be encoded using a cylindrical constraint on the position of the marker:
\begin{equation}
        \begin{split}
    \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
        0 \leq p_{p_m,z} &\leq h
        \end{split}
\end{equation}
where $r$ is the radius of the cylinder and $h$ is the height of the cylinder.
The marker of course also cant be within the mesh of the manufacturing system, but since this would result in a line of sight visibility of 0 we do not need to encode this constraint explicitly.
If we still want the marker close to the surface of the system such as when mounting them magnetically we can add a penalty function to the objective function that penalizes markers that are too far away from the flange.
This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.
